---
title: "Examining the draft success of NFL teams using Pro Football Reference's Approximate Value statistic"
author: "Myron Keith Gibert Jr"
date: "March 4, 2020"
output: pdf_document
toc: true
---
```{r tinytex, include=FALSE, results = "hide",warning=FALSE}
## INSTALLING LATEX FOR RMARKDOWN

#RMarkdown requires LaTex to create pdf documents as an output. More information can be found [here](https://bookdown.org/yihui/rmarkdown/pdf-document.html). Alternatively, output can be set to "html_document" or "word_document". End users may install LaTex by setting the "wanttinytex" variable to FALSE and running the following script:

#Install tinytex to let RMarkdown create a pdf document? Default: wanttinytex <- FALSE
wanttinytex <- FALSE

if(tinytex:::is_tinytex()==FALSE && wanttinytex == TRUE){
if (!require("tinytex")) install.packages("tinytex")
tinytex::install_tinytex()
}
```


## Introduction
Examining the draft success of NFL teams using Pro Football Reference's Approximate Value statistic. 

Moved to Google Drive.

https://journals.sagepub.com/doi/full/10.1177/0569434518812678

https://fisherpub.sjfc.edu/cgi/viewcontent.cgi?article=1064&context=sport_undergrad

https://repository.upenn.edu/cgi/viewcontent.cgi?article=1552&context=statistics_papers

https://www.forbes.com/sites/prishe/2015/05/22/tracking-nfl-draft-efficiency-how-contingent-is-success-to-draft-position/?sh=5edcca0f7495

https://www.pff.com/news/nfl-investigating-the-individual-draft-success-of-every-nfl-team

https://digitalcommons.bryant.edu/cgi/viewcontent.cgi?article=1012&context=honors_economics

https://creativematter.skidmore.edu/cgi/viewcontent.cgi?article=1040&context=econ_studt_schol

https://www.providencejournal.com/article/20130924/sports/309249910

http://www.footballperspective.com/draft-value-chart/

http://www.footballperspective.com/creating-a-draft-value-chart-part-ii/

## Set parameters

This initial section is used to set the dependent variables that are used for this analysis. 

THIS IS THE ONLY SECTION THAT SHOULD BE MODIFIED BY END USERS. 

The following variables can be set:

* outputdir is a character variable containing the "output directory", or the location where all output files will go. This can be any characters enclosed in quotations marks (ex. "Outputs"). Set to "Outputs" to use the default outputs folder
* ago1, ago2, and ago3 are logical variable that should only be set to TRUE or FALSE. This determines which of the PAR-CLiP data files for each argonaute protein will be used. Set each to TRUE to include all PAR-CLiP data. 
* tcga.fdr is a numerical variable that is used as a p-value cutoff for the fold change (as determined via DESeq2). A typical value is 0.05 or 0.01. This means that only fold changes with an multiple hypothesis adjusted p-value that is less that or equal to this variable will be included in the analysis.  Set to 1 to include all genes.
* cox.cutoff, like tcga.fdr, is a numerical variable that is used as a cutoff for the cox survival coefficient.  Any values that are >= the cutoff will be included in the analysis.  The default is a cox coefficient of 0.3. A number closer to zero is considered more weakly correlated and a number closer to 1 is considered more strongly correlated.
* deleteoutputs is a logical vector that should only be set to TURE or FALSE.  When set to true, the output directory will be repopulated after every program run.  If set to false, the program will only run if the output directory does not already exist in the working directory, to prevent the data from being overwritten.

```{r parameters}

outputdir <- "Outputs"

currentyear <- 2021

#years <- c(2010,2011)
years <- c(2010,2011,2012,2013,2014,2015,2016,2017,2018,2019,2020)

deletespec <- TRUE

```

## Setup

This initializes all programs that will be used for the analysis.  This program will also auto-install any required packages that are not already installed. Lastly, this section outlines the parameters for knitr files. Most of the analysis is performed using the Tidyverse family of packages.  More information can be found at [tidyverse.org/packages](tidyverse.org/packages).  The readxl and xlsx packages are used for reading and writing .xlsx Excel workbooks, respectively.  

```{r setup, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

if (!require("htm2txt")) install.packages("htm2txt")
library(htm2txt)

if (!require("stringr")) install.packages("stringr")
library("stringr")

if (!require("stringi")) install.packages("stringi")
library("stringi")

if (!require("splitstackshape")) install.packages("splitstackshape")
library("splitstackshape")

if (!require("tidyverse")) install.packages("tidyverse")
library("tidyverse")

if (!require("ggplot2")) install.packages("ggplot2")
library("ggplot2")

if (!require("ggforce")) install.packages("ggforce")
library("ggforce")

if (!require("ggrepel")) install.packages("ggrepel")
library("ggrepel")

if (!require("ggfortify")) install.packages("ggfortify")
library("ggfortify")

```

## Setup

Create output directory.

```{r outputs}

if(!dir.exists(outputdir)){
  dir.create(outputdir)
}

```

## Setup

Create output directory.

## Create mastertable

```{r textscrape, message = FALSE, warning = FALSE}

mastertable <- data.frame()

i <- 1

for(i in 1:length(years)){
print(paste("Adding data for ",years[i]," Draft to master table...",sep=""))

url <- paste("https://www.pro-football-reference.com/years/",years[i],"/draft.htm#drafts",sep="")

download.file(url, "nfldraft.csv", method="curl")

draftdata <- as.data.frame(readLines("nfldraft.csv"))

#draftdata$Rnd <- sapply(draftdata$`readLines("nfldraft.csv")`,function(x) str_extract(x,'(?<=STRING1).*(?=STRING2)')[1])

draftdata$Rnd <- sapply(draftdata$`readLines("nfldraft.csv")`,function(x) str_extract(x,'(?<=data-stat="draft_round" >).*(?=</th><td class=)')[1])

draftdata$Pick <- sapply(draftdata$`readLines("nfldraft.csv")`,function(x) stri_extract_first_regex(str_extract(x,'(?<=5" >).*(?=</td><td class=)'),"[0-9]+")[1])

draftdata$Tm <- sapply(draftdata$`readLines("nfldraft.csv")`,function(x) substr(str_extract(x,'(?<=">).*(?=</a></td><td class="left ")'),1,3)[1])
mastertable$Tm <- ifelse(mastertable$Tm == "STL","LAR",ifelse(mastertable$Tm == "SDG","LAC",ifelse(mastertable$Tm == "OAK","LVR",as.character(mastertable$Tm))))

draftdata$Player <- sapply(draftdata$`readLines("nfldraft.csv")`,function(x) str_extract(str_extract(x,'(?<=.htm">).*(?=</a>)'), "[^<]+")[1])

draftdata$Player2 <- sapply(draftdata$`readLines("nfldraft.csv")`,function(x) (str_extract(str_extract(x,'(?<=\"player\" csk=\").*(?=\" >)'), '[^"]+')[1]))

draftdata <- draftdata %>% 
    separate(Player2, c("First", "Last"), remove = F)

draftdata$Player <- ifelse(is.na(draftdata$Player),paste(draftdata$First,draftdata$Last,sep=" "),draftdata$Player)

draftdata$Pos <- sapply(draftdata$`readLines("nfldraft.csv")`,function(x) str_remove_all(str_extract(x,'(?<=data-stat=\"pos\" csk=\").*(?=---)'),"-")[1])

draftdata$year <- years[i]

draftdata$CarAV <- sapply(draftdata$`readLines("nfldraft.csv")`,function(x) substr(str_extract(x,'(?<=data-stat=\"career_av\" >).*(?=</td><td class=)'),1,3)[1])
                        
draftdata$CarAV <- as.numeric(gsub("([0-9]+).*$", "\\1", draftdata$CarAV))

draftdata$CarAV[is.na(draftdata$CarAV)] <- 0

draftdata$DrAV <- sapply(draftdata$`readLines("nfldraft.csv")`,function(x) substr(str_extract(x,'(?<=data-stat=\"draft_av\" >).*(?=</td><td class=)'),1,3)[1])
                        
draftdata$DrAV <- as.numeric(gsub("([0-9]+).*$", "\\1", draftdata$DrAV))

draftdata$DrAV[is.na(draftdata$DrAV)] <- 0

draftdata_complete <- as.data.frame(draftdata[complete.cases(draftdata),-1])

draftdata_complete <- draftdata_complete %>%
  select(-First,-Last,-Player2)

mastertable <- rbind(mastertable,draftdata_complete)

}

mastertable <- mastertable %>%
  mutate(Pick =ifelse(
    (Rnd == 1 & Pick == 1),1,ifelse(
    (Rnd == 2 & Pick == 1),64,ifelse(
    (Rnd == 3 & Pick == 1),96,ifelse(
    (Rnd == 4 & Pick == 1),128,ifelse(
    (Rnd == 5 & Pick == 1),160,ifelse(
    (Rnd == 6 & Pick == 1),196,ifelse(
    (Rnd == 7 & Pick == 1),228,Pick
  ))))))))

write.csv(mastertable,"mastertable.csv")

```

## Create recordtable

```{r recordtable}

recordtable <- data.frame()

j <- 2

for(j in 1:length(years)){
print(paste("Adding data for ",years[j]," to record table...",sep=""))

  
recordurl <- paste("https://www.pro-football-reference.com/years/",years[j],"/",sep="")
  
download.file(recordurl, "nfcrecord.csv", method="curl")

nfctext <- as.data.frame(readLines("nfcrecord.csv"))
colnames(nfctext) <- "Text"

##recorddate scraper

recorddata <- data.frame()
  
nfctext$Tm <- sapply(nfctext$Text,function(x) substr(str_extract(x,'(?<=<a href="/teams/).*(?=/)'),1,3)[1])

nfctext$Tm <- ifelse((nfctext$Tm == "nyj"),"NYJ",ifelse(
    (nfctext$Tm == "mia"),"MIA",ifelse(
    (nfctext$Tm == "nwe"),"NWE",ifelse(
    (nfctext$Tm == "buf"),"BUF",ifelse(
    (nfctext$Tm == "pit"),"PIT",ifelse(
    (nfctext$Tm == "rav"),"BAL",ifelse(
    (nfctext$Tm == "cle"),"CLE",ifelse(
    (nfctext$Tm == "cin"),"CIN",ifelse(
    (nfctext$Tm == "clt"),"IND",ifelse(
    (nfctext$Tm == "jax"),"JAX",ifelse(
    (nfctext$Tm == "htx"),"HOU",ifelse(
    (nfctext$Tm == "oti"),"TEN",ifelse(
    (nfctext$Tm == "kan"),"KAN",ifelse(
    (nfctext$Tm == "sdg"),"LAC",ifelse(
    (nfctext$Tm == "rai"),"LVR",ifelse(
    (nfctext$Tm == "den"),"DEN",ifelse(
    (nfctext$Tm == "phi"),"PHI",ifelse(
    (nfctext$Tm == "nyg"),"NYG",ifelse(
    (nfctext$Tm == "dal"),"DAL",ifelse(
    (nfctext$Tm == "was"),"WAS",ifelse(
    (nfctext$Tm == "chi"),"CHI",ifelse(
    (nfctext$Tm == "gnb"),"GNB",ifelse(
    (nfctext$Tm == "det"),"DET",ifelse(
    (nfctext$Tm == "min"),"MIN",ifelse(
    (nfctext$Tm == "atl"),"ATL",ifelse(
    (nfctext$Tm == "nor"),"NOR",ifelse(
    (nfctext$Tm == "tam"),"TAM",ifelse(
    (nfctext$Tm == "car"),"CAR",ifelse(
    (nfctext$Tm == "sea"),"SEA",ifelse(
    (nfctext$Tm == "ram"),"LAR",ifelse(
    (nfctext$Tm == "sfo"),"SFO",ifelse(
    (nfctext$Tm == "crd"),"ARI",ifelse("ERROR")
    ))))))))))))))))))))))))))))))))

nfctext$year <- years[j]
  
nfctext$Wins <- sapply(nfctext$Text,function(x) substr(str_extract(x,'(?<= data-stat="wins" >).*(?=</td><td class="right " )'),1,2)[1])
nfctext$Wins <- as.numeric(gsub("([0-9]+).*$", "\\1", nfctext$Wins))

nfctext$Losses <- sapply(nfctext$Text,function(x) substr(str_extract(x,'(?<= data-stat="losses" >).*(?=</td><td class="right)'),1,2)[1])
nfctext$Losses <- as.numeric(gsub("([0-9]+).*$", "\\1", nfctext$Losses))

nfctext$Ties <- sapply(nfctext$Text,function(x) substr(str_extract(x,'(?<=data-stat="ties" >).*(?=</td><td class="right)'),1,2)[1])
nfctext$Ties <- as.numeric(gsub("([0-9]+).*$", "\\1", nfctext$Ties))

nfctext$PointDif <- sapply(nfctext$Text,function(x) substr(str_extract(x,'(?<=data-stat="points_diff" >).*(?=</td><td class="right)'),1,3)[1])
nfctext$PointDif <- as.numeric(gsub("([0-9]+).*$", "\\1", nfctext$PointDif))

nfctext$SRS <- sapply(nfctext$Text,function(x) substr(str_extract(x,'(?<=data-stat="srs_total" >).*(?=</td><td class)'),1,3)[1])
nfctext$SRS <- as.numeric(gsub("([0-9]+).*$", "\\1", nfctext$SRS))

recorddata_complete <- as.data.frame(nfctext[complete.cases(nfctext),-1])

recorddata_complete <- recorddata_complete %>%
  mutate_all(~replace(., ( is.na(.) | . == ""), 0)) 

recordtable <- rbind(recordtable,recorddata_complete)

recordtable <- unique(recordtable)

}

recordtable <- recordtable %>%
  select(-year)

#recordtable$Wins <- as.numeric(recordtable$Wins)
#recordtable$PointDif <- as.numeric(recordtable$PointDif)
#recordtable$SRS <- as.numeric(recordtable$SRS)
write.csv(recordtable,"recordtable.csv")
recordtable <- read.csv("recordtable.csv",header=TRUE,row.names = 1)
recordtable_sum <- recordtable %>% 
  unique() %>%
  group_by(Tm) %>%
  summarise(TotalWins=sum(Wins,na.rm = TRUE),TotalNetSRS=sum(SRS)/length(years),TotalNetPointDif=sum(PointDif)/length(years))

write.csv(recordtable_sum,"recordtable.csv")

```

# Create some functions

```{r functions}

NFL_calcdif <- function(mastertable, valuechart,recordtable){
#mastertable <- mt_perm #read.csv("mastertable.csv", header = TRUE)
#recordtable <- read.csv("recordtable.csv",header = TRUE)
#valuechart <- valuechart
mastertable <- mastertable %>% left_join(valuechart,by="Pick")
mastertable <- mastertable %>% left_join(recordtable,by="Tm")

mastertable <- mastertable %>% 
  mutate(year_dif = currentyear - year) %>%
  mutate(ExAV.adj = ifelse(year_dif >= 4,ExAV,ExAV/(4-year_dif))) %>%
  mutate(dif = DrAV - ExAV.adj, 
         hit = ifelse(dif>=0,1,0),
         gotaway.dif = CarAV - DrAV,
         gotaway = ifelse(DrAV<gotaway.dif,1,0)) 
mastertable <- mastertable %>% group_by(Tm,year) %>% mutate(class.dif = sum(dif,na.rm=TRUE)) 
#mt_calc <- mastertable
mastertable
}

NFL_summarize_Tm <- function(mastertable = mastertable,valuechart = valuechart) {
  #mastertable <- mt_calc
  #mastertable[is.na(mastertable)] <- 0
  mastertable_sum <- mastertable %>%
  group_by(Tm) %>%
  mutate(picks = n(),Netvalue = sum(dif,na.rm=TRUE),TotalNetAVLost = sum(gotaway.dif,na.rm=TRUE),PickEfficiency = (Netvalue-TotalNetAVLost)/picks) %>%
  select(Tm,TotalWins,TotalNetSRS,TotalNetPointDif,picks,Netvalue,TotalNetAVLost,PickEfficiency) %>%
  #summarise(hits = sum(hit,na.rm=TRUE),picks = length(hit),hit.percentage = hits/picks,totalvalue = sum(dif,na.rm=TRUE)) %>%
  ungroup() %>%
  unique()
}

NFL_permute <- function(mastertable = mastertable,valuechart = valuechart,recordtable = recordtable,nperms = nperms,seed_number = 20210120){

#mastertable <- read.csv("mastertable.csv", header = TRUE)
#valuechart <- valuechart
#nperms <- 100
#seed_number <- 20210120

set.seed(seed_number)
  
mt_sum <- data.frame()

mt_calc <- NFL_calcdif(mastertable,valuechart,recordtable)

realvalue <- NFL_summarize_Tm(mt_calc)
colnames(realvalue) <- c("Tm","TotalWins","TotalNetSRS","TotalNetPointDif","picks","RealNetValue","RealTotalNetAVLost","RealPickEfficiency")
  
real_picks <- mastertable %>%
  select(Tm,Rnd,Pick,year)

i <- 1

#for(i in 1:10){
for(i in 1:nperms){
 
print(i)
perms <- mastertable %>%
  group_by(Rnd,year) %>%
  sample_frac(1) %>% 
  ungroup() %>%
  group_by(year) %>%
  mutate(pick = row_number()) %>%
  select(-Pick) %>%
  mutate("Pick" = pick) %>%
  select(-pick,-X)

mt_perm <- real_picks %>% 
  left_join(perms, by = c("Pick","year")) %>%
  select(-Tm.y,"Tm"=Tm.x)

mt_bind <- NFL_calcdif(mt_perm,valuechart,recordtable)
mt_bind <- NFL_summarize_Tm(mt_bind)
mt_sum <- rbind(mt_sum,mt_bind)
}

mt_final <- mt_sum %>%
  left_join(realvalue,by="Tm") %>%
  mutate(p.dif = as.numeric(RealPickEfficiency) - as.numeric(PickEfficiency)) %>%
  mutate(p.cases = ifelse(p.dif<0,1,0)) %>%
  group_by(Tm) %>%
  mutate("TotalNetDrAV" = round(RealNetValue),p = sum(p.cases)/nperms) %>%
  #mutate(padj = p.adjust(p,method="fdr")) %>%
  ungroup() %>%
  select(Tm,"TotalWins"=TotalWins.y,"NetPointDif"=TotalNetPointDif.y,"NetSRS"=TotalNetSRS.y,"picks"=picks.y,TotalNetDrAV,"TotalNetAVLost" = RealTotalNetAVLost,"PickEfficiency"=RealPickEfficiency,p) %>%
  unique() 

}

```

```{r summary}

all_content <- readLines("./draftvaluechart.csv")
skip_lines <- all_content[-1]
valuechart <- read.csv(textConnection(skip_lines), header = TRUE, stringsAsFactors = FALSE)
valuechart <- valuechart %>% select(Pick,"ExAV" = AV)

mastertable <- read.csv("mastertable.csv", header = TRUE)

recordtable <- read.csv("recordtable.csv",header = TRUE)

teams <- unique(mastertable$Tm)

summarytable <- NFL_summarize_Tm(mt_calc)

summarytable <- NFL_permute(mastertable,valuechart,recordtable = recordtable,nperms = 10000)

```

```{r bestvalueplayers}

bestvalueplayers <- mt_calc %>% 
             group_by(Tm) %>%
             filter(dif == max(dif,na.rm=TRUE)) %>%
             select(year,Rnd,Pick,Tm,Player,Pos,ExAV,DrAV,dif)

```

```{r bestplayers}

bestplayers <- mt_calc %>% 
             group_by(Tm) %>%
             filter(DrAV == max(DrAV,na.rm=TRUE)) %>%
             select(year,Rnd,Pick,Tm,Player,Pos,ExAV,DrAV,dif)

```

```{r biggestbusts}

biggestbusts <- mt_calc %>% 
             group_by(Tm) %>%
             filter(year <= 2016) %>%
             filter(dif == min(dif,na.rm=TRUE)) %>%
             select(year,Rnd,Pick,Tm,Player,Pos,ExAV,DrAV,dif)

```

```{r eaglesplayers}

eaglesplayers <- mt_calc %>% 
             filter(Tm == "PHI") %>%
             select(year,Rnd,Pick,Tm,Player,Pos,ExAV,DrAV,dif)

seahawksplayers <- mt_calc %>% 
             filter(Tm == "SEA") %>%
             select(year,Rnd,Pick,Tm,Player,Pos,ExAV,DrAV,dif)

```

```{r}

TheOneThatGotAway <- mt_calc %>% 
  group_by(Tm) %>%
  mutate(TotalNetAVLost = sum(gotaway.dif)) %>%
  filter(gotaway.dif == max(gotaway.dif)) %>%
  select(Tm,Player,Pos,year,CarAV,DrAV,"NetAVLost" = gotaway.dif,TotalNetAVLost)

```

```{r}

bestpositions <- mt_calc %>% group_by(Pos) %>%
  summarize(medianExAV = median(ExAV,na.rm=TRUE), medianDrAV = median(DrAV,na.rm=TRUE), medianDiff = median(dif,na.rm=TRUE),nplayers = length(dif))

```

```{r safestposition}

safestpicksRnd1 <- mt_calc %>% 
             filter(Rnd == 1) %>%
             group_by(Pos) %>%
  summarise(totalvalue = sum(dif,na.rm=TRUE), npicks = n(), averagevalue = sum(dif,na.rm=TRUE)/n())

safestpicksRnd2 <- mt_calc %>% 
             filter(Rnd == 2) %>%
             group_by(Pos) %>%
  summarise(totalvalue = sum(dif,na.rm=TRUE), npicks = n(), averagevalue = sum(dif,na.rm=TRUE)/n())

safestpicksRnd3 <- mt_calc %>% 
             filter(Rnd == 3) %>%
             group_by(Pos) %>%
  summarise(totalvalue = sum(dif,na.rm=TRUE), npicks = n(), averagevalue = sum(dif,na.rm=TRUE)/n())

safestpicksRnd4 <- mt_calc %>% 
             filter(Rnd == 4) %>%
             group_by(Pos) %>%
  summarise(totalvalue = sum(dif,na.rm=TRUE), npicks = n(), averagevalue = sum(dif,na.rm=TRUE)/n())

safestpicksRnd5 <- mt_calc %>% 
             filter(Rnd == 5) %>%
             group_by(Pos) %>%
  summarise(totalvalue = sum(dif,na.rm=TRUE), npicks = n(), averagevalue = sum(dif,na.rm=TRUE)/n())

safestpicksRnd6 <- mt_calc %>% 
             filter(Rnd == 6) %>%
             group_by(Pos) %>%
  summarise(totalvalue = sum(dif,na.rm=TRUE), npicks = n(), averagevalue = sum(dif,na.rm=TRUE)/n())

safestpicksRnd7 <- mt_calc %>% 
             filter(Rnd == 7) %>%
             group_by(Pos) %>%
  summarise(totalvalue = sum(dif,na.rm=TRUE), npicks = n(), averagevalue = sum(dif,na.rm=TRUE)/n())

safestpicksall <- mt_calc %>% 
             group_by(Pos) %>%
  summarise(totalvalue = sum(dif,na.rm=TRUE), npicks = n(), averagevalue = sum(dif,na.rm=TRUE)/n())

```

```{r}

if(deletespec == TRUE){unlink(c("nfldraft.csv","nfcrecord.csv","mastertable.csv","recordtable.csv"),recursive = TRUE)}

```

